cardinalR: Generating interesting high-dimensional data structures

### 2025-11-08 paper-cardinalR.Rmd
Trying to make the paper shorter

```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow for generating high-dimensional clustered data. The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise). Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset."}

@@

```{r, echo=TRUE}
# Define example rotation matrices for 4D space
rot1 <- gen_rotation(p = 4, planes_angles = list(
  list(plane = c(1, 2), angle = 60), list(plane = c(3, 4), angle = 90)
))

# Generate the clustered dataset
clust_data <- gen_multicluster(
  n = c(200, 300, 500),
  k = 3,
  loc = matrix(c(
    0, 0, 0, 0,
    5, 9, 0, 0,
    3, 4, 10, 7
  ), nrow = 3, byrow = TRUE),
  scale = c(3, 1, 2),
  shape = c("gaussian", "cone", "unifcube"),
  rotation = list(rot1, NULL, NULL),
  is_bkg = FALSE
)

@@

@@

## To generate data
five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
```

@@

The included structures cover a wide range of diagnostic settings. Branching shapes facilitate the study of continuity and topological preservation, the Scurve with a hole allows investigation of incomplete manifolds, and clustered spheres assess separability on curved surfaces. The Möbius strip introduces challenges from non-orientable geometry, while gridded cubes and pyrholes test spatial regularity and clustering in sparse, non-convex regions.

>>>>

```{r workflow, out.width = "70%", fig.pos="H", fig.align='center', fig.cap="Workflow for generating high-dimensional clustered data. The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise). Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset."}

@@

```{r, echo=TRUE}
rot1 <- gen_rotation(p = 4, planes_angles = list(
  list(plane = c(1, 2), angle = 60), list(plane = c(3, 4), angle = 90)
))

clust_data <- gen_multicluster(
  n = c(200, 300, 500),
  k = 3,
  loc = matrix(c(
    0, 0, 0, 0,
    5, 9, 0, 0,
    3, 4, 10, 7
  ), nrow = 3, byrow = TRUE),
  scale = c(3, 1, 2),
  shape = c("gaussian", "cone", "unifcube"),
  rotation = list(rot1, NULL, NULL),
  is_bkg = FALSE
)

@@

<!-- $$ -->
<!-- X_i = -->
<!-- \begin{cases} -->
<!-- \displaystyle \prod_{j=1}^{i-1} \sin(\theta_j) , \cos(\theta_i), & 1 \le i \le p-1, \\ -->
<!-- \displaystyle \prod_{j=1}^{p-1} \sin(\theta_j), & i = p, -->
<!-- \end{cases} -->
<!-- $$ -->

<!-- with the convention that $\prod_{j=1}^{0} \sin(\theta_j) = 1$. -->

@@

five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
```

@@

The included structures cover a wide range of diagnostic settings. Branching shapes facilitate the study of continuity and topological preservation, the S-curve with a hole allows investigation of incomplete manifolds, and clustered spheres assess separability on curved surfaces. The Möbius strip introduces challenges from non-orientable geometry, while gridded cubes and pyrholes test spatial regularity and clustering in sparse, non-convex regions.

### 2025-11-08 paper-cardinalR.tex
Trying to make the paper shorter

\begin{figure}[H]
\includegraphics[width=1\linewidth]{figures/cardinalR_workflow} \caption{Workflow for generating high-dimensional clustered data. The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise). Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.}\label{fig:workflow}
\end{figure}

@@

\begin{verbatim}
# Define example rotation matrices for 4D space
rot1 <- gen_rotation(p = 4, planes_angles = list(
  list(plane = c(1, 2), angle = 60), list(plane = c(3, 4), angle = 90)
))

# Generate the clustered dataset
clust_data <- gen_multicluster(
  n = c(200, 300, 500),
  k = 3,
  loc = matrix(c(
    0, 0, 0, 0,
    5, 9, 0, 0,
    3, 4, 10, 7
  ), nrow = 3, byrow = TRUE),
  scale = c(3, 1, 2),
  shape = c("gaussian", "cone", "unifcube"),
  rotation = list(rot1, NULL, NULL),
  is_bkg = FALSE
)
\end{verbatim}

@@

## To generate data
five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
\end{verbatim}

>>>>

{\centering \includegraphics[width=0.7\linewidth]{figures/cardinalR_workflow}

\caption{Workflow for generating high-dimensional clustered data. The user specifies input parameters (number of points, clusters, cluster shapes, scaling, rotation, and optional background noise). Clusters are generated iteratively, transformed, optionally augmented with Gaussian noise dimensions, combined, and labeled, resulting in the final dataset.}\label{fig:workflow}
\end{figure}

@@

\begin{verbatim}
rot1 <- gen_rotation(p = 4, planes_angles = list(
  list(plane = c(1, 2), angle = 60), list(plane = c(3, 4), angle = 90)
))

clust_data <- gen_multicluster(
  n = c(200, 300, 500),
  k = 3,
  loc = matrix(c(
    0, 0, 0, 0,
    5, 9, 0, 0,
    3, 4, 10, 7
  ), nrow = 3, byrow = TRUE),
  scale = c(3, 1, 2),
  shape = c("gaussian", "cone", "unifcube"),
  rotation = list(rot1, NULL, NULL),
  is_bkg = FALSE
)
\end{verbatim}

@@

five_clusts <- gen_multicluster(n = c(2250, 1500, 750, 1250, 1750), k = 5,
                       loc = positions,
                       scale = c(0.25, 0.35, 0.3, 1, 0.3),
                       shape = c("helicalspiral", "hemisphere", "unifcube", 
                                 "cone", "gaussian"),
                       rotation = NULL,
                       is_bkg = FALSE)
\end{verbatim}
